Spin canonical divisors of log stacky curves

### 2015-06-29 stacky_write_up/spin_cring.tex
main corollary revision, minor typographical things

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bO{{\mathbf O}}
\newcommand\CC{{\mathcal C}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BBZ{{\mathbb Z}}
\newcommand\uR{\underline{R}}
\newcommand\uZ{\underline{\BBZ}}
\newcommand\CF{{\mathcal F}}
\newcommand\uCF{\underline{{\mathcal F}}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}
\newcommand \sx{\mathscr X}

@@

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\bh$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, we sharply bound the degree of the generators and
relations for $M(\Gamma)$, for general $\Gamma$ \todo{Peter: Why do you say general gamma?  I think you should either add "general fuchsian groups gamma" or just take out that clause, since you just defined what gamma is.}  \todo{Peter: also, its not clear to me that this should be said right here.  I think we may want to say a bit more about what's already known instead of going directly to what we have done}.  In the special that $\Gamma$ has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give similar bounds, and we extend their work to all Fuchsian groups $\Gamma$. In order to find generators and relations for $M(\Gamma)$, we translate the seemingly analytic question of understanding the ring of modular forms into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$
. Hence, for the remainder of the paper, we will work in the
algebraic category.

@@

\noindent
\todo{incorporate some chunk on O'Dorney somewhere in intro, below is previously written stuff}
In particular, we would like to bound the degrees in which the
graded ring and its relations are generated. The work by O'Dorney
shows that the canonical ring of any $\BQ$-divisor $D =
\frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of
$\BP_k^1$ with $\# \Supp(D) \leq 2$ is generated in degrees up
to $e := \max(\beta_1, \beta_2)$ with relations generated in
degrees up to $2e$ \cite[Section 4]{dorney:canonical}. Furthermore,
O'Dorney gives a larger bound when the support of $D$ is allowed to
have arbitrarily many points in terms of least-common-multiples.
Voight and Zureick-Brown also bounded the generator and relation
degrees of log spin canonical rings of $(\sx, \Delta, \halfcan)$,
in the case that genus is positive and $\halfcan$ is effective,
by $3e$ and $6e$ for $e := \max(e_1, \ldots, e_r)$.

@@

%uses best lower approximations to provide a minimal presentation for the canonical ring of an arbitrary $\BQ$-divisor $D = \frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of $\BP_k^1$ with $\# \Supp(D) \leq 2$ and shows that it is generated in degree $e := \max(\beta_1, \beta_2)$ with relations generated in degrees at most $2e$ \cite[Section 4]{dorney:canonical}. For arbitrarily many points, O'Dorney also provides much larger bounds \cite[Section 5]{dorney:canonical}. While
%
%In Lemma ~\ref{lem:sat-1} we provide an inductive method to strengthen
%O'Dorney's result \cite[Theorem 8]{dorney:canonical} on $\BP^1_k$ to divisors $D = \sum_{i = 1}^n \frac{\alpha_i
%}{\beta_i} P_i$ for points $P_1, \ldots, P_n$ on a curve $X$ such that $H^
%0(X, \lfloor D \rfloor) > 0$. Our result implies that $R_D$ is generated in
%degrees at most $e = \max(\beta_i)$ with relations generated in
%degrees at most $2e$. \todo{Are you sure you're not assuming $\frac{\alpha}{\beta}$ is effective? If not, doesn't what you say contradict Evan's example showing that his bounds of about lcm are asymptotically sharp?} Furthermore, given additional constraints on $H^0(
%X, \lfloor dD \rfloor)$ for various values of $d$, we can apply the
%same Riemann-Roch \todo{same as what? Maybe just say we can generalize his results to arbitrary genus?} arguments in arbitrary genus. 
%To do this, we extend a result by Voight and
%Zureick-Brown \cite[Theorem 8.3.1]{vzb:stacky} on stacky canonical rings to arbitrary $\BQ$-
%divisors. \todo{Again, I don't think it can be arbitrary, and what are the $\BQ$ divisors on?}

@@

\begin{cor}
\label{cor:main-mod-forms}
There exists a log half-canonical divisor $\halfcan$ on $\sx$ such
that we have a graded isomorphism of $\BC$-algebras
\[
	\bigoplus_{k = 0}^{\infty} M_k(\Gamma) \cong R(\sx (\Gamma), 
\Delta, \halfcan)
\]

\noindent
Let $\sigma = (g; e_1, \ldots, e_r; \delta)$ be the signature
of $\sx$. Then $\bigoplus_{k = 0}^{\infty} M_k(\Gamma)$ is
generated as a $\BC$-algebra by elements of degree at most $e = \max
(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$
then $R(\sx,\Delta, \halfcan)$ has relations generated in degrees
up to $\max(11, 2e_1, \ldots, 2e_r)$. Otherwise, $R(\sx, \Delta,
\halfcan)$ has relations generated in degrees up to $\max(10, 2e_1,
\ldots, 2e_r),$ so long as $\sigma$ does not lie in a finite list
of exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional}
for signatures with $g = 0$.
\end{cor}

@@

\begin{defn}
The {\bf log spin canonical ring} of $(\sx, \Delta, \halfcan)$ is
\[
	R(\sx, \Delta, L) = \bigoplus_{d \geq 0} H^0(\sx, \lfloor d L \rfloor)
\]
\end{defn}
\todo{what about floors?}

@@

We can then consider the
orbifold $\Gamma \backslash \bh$ over $\BC$. For example, in the
case $\Gamma$ acts freely on $\bh$, $\Gamma \backslash \bh$ is a
Riemann Surface over $\BC$. It may well be that $\Gamma \backslash
\bh$ is not compact. Defining $\Delta$ as the divisor of cusps
associated to $\Gamma$, we can form a space $\Gamma \backslash
\bh^*$ which is a compactification of $\Gamma \backslash \bh,$ with
associated {\bf divisor of cusps} $\Delta$.

>>>>

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\BH{{\mathbb H}}
\newcommand\bO{{\mathbf O}}
\newcommand\CC{{\mathcal C}}
\newcommand\BN{{\mathbb N}}
\newcommand\BC{{\mathbb C}}
\newcommand\BF{{\mathbb F}}
\newcommand\BR{{\mathbb R}}
\newcommand\BQ{{\mathbb Q}}
\newcommand\BP{{\mathbb P}}
\newcommand\BBZ{{\mathbb Z}}
\newcommand\uR{\underline{R}}
\newcommand\uZ{\underline{\BBZ}}
\newcommand\CF{{\mathcal F}}
\newcommand\uCF{\underline{{\mathcal F}}}
\newcommand\BZ{{\mathbb Z}}
\newcommand\BA{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}
\newcommand \sx{\mathscr X}

@@

\section{Introduction}
Let $\Gamma$ be a {\bf Fuchsian} group, i.e. a discrete subgroup
of $PSL_2(\BR)$ acting on the upper half plane $\BH$ by fractional
linear transformations. We consider the {\bf ring of modular forms}
$M(\Gamma) = \bigoplus_k M_k(\Gamma)$. In this paper, we sharply bound the degree of the generators and
relations for $M(\Gamma)$, for general $\Gamma$ \todo{Peter: Why do you say general gamma?  I think you should either add "general fuchsian groups gamma" or just take out that clause, since you just defined what gamma is.}  \todo{Peter: also, its not clear to me that this should be said right here.  I think we may want to say a bit more about what's already known instead of going directly to what we have done}.  In the special that $\Gamma$ has no odd weight modular forms, Voight and Zureick-Brown
\cite[Chapters 7-9]{vzb:stacky} give similar bounds, and we extend their work to all Fuchsian groups $\Gamma$. In order to find generators and relations for $M(\Gamma)$, we translate the seemingly analytic question of understanding the ring of modular forms into the algebraic
category, using a generalization of the GAGA principle. By
\cite[Proposition 6.1.5]{vzb:stacky}, there is an equivalence of
categories between orbifold curves and log stacky curves over $\BC$
. Hence, for the remainder of the paper, we will work in the
algebraic category.

@@

In particular, we would like to bound the degrees in which the
graded ring and its relations are generated. The work by O'Dorney
shows that the canonical ring of any $\BQ$-divisor $D =
\frac{\alpha_1}{\beta_1}P_1 + \frac{\alpha_2}{\beta_2}P_2$ of
$\BP_k^1$ with $\# \Supp(D) \leq 2$ is generated in degrees up
to $e := \max(\beta_1, \beta_2)$ with relations generated in
degrees up to $2e$ \cite[Section 4]{dorney:canonical}. Furthermore,
O'Dorney gives a larger bound when the support of $D$ is allowed to
have arbitrarily many points in terms of least-common-multiples.
Voight and Zureick-Brown also bounded the generator and relation
degrees of log spin canonical rings of $(\sx, \Delta, \halfcan)$,
in the case that genus is positive and $\halfcan$ is effective,
by $3e$ and $6e$ for $e := \max(e_1, \ldots, e_r)$.

@@

@@

\begin{cor}
\label{cor:main-mod-forms}
Let $\Gamma$ be a Fuchsian group and $\sx$ the stacky curve
associated to $\Gamma \backslash \BH$ with signature $(g; e_1,
\ldots, e_r; \delta)$. Then the ring of modular forms
$M(\Gamma) := \bigoplus_{k = 0}^{\infty} M_k(\Gamma)$ is
generated as a $\BC$-algebra by elements of degree at most $e = \max
 (5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$
then $M(\Gamma)$ has relations generated in degrees up to $\max(11,
2e_1, \ldots, 2e_r)$. Otherwise, $M(\Gamma)$ has relations
generated in degrees up to $\max(10, 2e_1, \ldots, 2e_r)$, so long
as $\sigma$ does not lie in a finite list of exceptional cases, as
listed in Table ~\ref{table:g-1-exceptional} for signatures with $g
= 1$ and Table ~\ref{table:g-0-exceptional} for signatures with $g =
0$.
\end{cor}

@@

\begin{defn}
If divisor $D \in \di \sx$ and $D = \sum_{i = 1}^{n} \alpha_i P_i$
with $\alpha_i \in \BQ$, the floor of a divisor $\lfloor D
\rfloor$ is defined to be $\lfloor D \rfloor := \sum_{i = 1}^{n}
\lfloor \alpha_i \rfloor P_i$.

The {\bf log spin canonical ring} of $(\sx, \Delta, \halfcan)$ is
\[
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx, \lfloor d \halfcan \rfloor)
\]
\end{defn}

@@

We can then consider the
orbifold $\Gamma \backslash \BH$ over $\BC$. For example, in the
case $\Gamma$ acts freely on $\BH$, $\Gamma \backslash \BH$ is a
Riemann Surface over $\BC$. It may well be that $\Gamma \backslash
\BH$ is not compact. Defining $\Delta$ as the divisor of cusps
associated to $\Gamma$, we can form a space $\Gamma \backslash
\BH^*$ which is a compactification of $\Gamma \backslash \BH,$ with
associated {\bf divisor of cusps} $\Delta$.
